---
title: "The Emmys"
author: "NearAndDistant"
date: "27/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

dir.create(here::here('Week 39 : Emmy Awards'))

```

#### Importing
```{r}
library(tidyverse)

emmy_raw <- 
read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-21/nominees.csv')

kaggle_emmy_raw <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/data_science_with_r/main/Tidy_Tuesday/2021/Week%2039%20%3A%20Emmy%20Awards/the_emmy_awards.csv")

```

```{r}

emmy_temp <- 
emmy_raw %>%
  mutate(category = str_extract(category, "[^-]+")) %>%
  mutate(category = str_to_title(category)) %>%
  filter(year > 2019) %>%
  select(year , category , nominee = title , company = distributor , type) %>%
  distinct() # quite a few double counting going on in this datset so distinct allows us to bypass

kaggle_temp <- 
kaggle_emmy_raw %>%
  mutate(win = if_else(win == TRUE, "Winner", "Nominee")) %>%
  select(year , category , nominee , company , type = win)

emmys <-
  rbind(emmy_temp , kaggle_temp) %>%
  mutate(category = str_to_title(category)) %>%
  drop_na(company)

```

#### Cleaning
```{r}

emmys_cln <-
emmys %>%
  mutate(company = if_else(str_detect(company , "youtube|YouTube"), "YouTube", company),
         company = if_else(str_detect(company , "ABC|abc"), "ABC", company),
         company = if_else(str_detect(company , "AMC"), "AMC", company),
         company = if_else(str_detect(company , "Amazon|Prime"), "Amazon", company),
         company = if_else(str_detect(company , "Apple"), "Apple", company),
         company = if_else(str_detect(company , "Disney"), "Disney", company),
         company = if_else(str_detect(company , "CTN|cartoonnetwork|Cartoon Network"), "Cartoon Network", company),
         company = if_else(str_detect(company , "CBS"), "CBS", company),
         company = if_else(str_detect(company , "ComedyCentral|Comedy Central"), "Comedy Central", company),
         company = if_else(str_detect(company , "Discovery|discovery"), "Discovery Chanel", company),
         company = if_else(str_detect(company , "Fox|FOX"), "FOX", company),
         company = if_else(str_detect(company , "FX"), "FX", company),
         company = if_else(str_detect(company , "HBO"), "HBO", company),
         company = if_else(str_detect(company , "NatGeo|Nat Geo|National Geographic"), "National Geographic", company),
         company = if_else(str_detect(company , "NBC"), "NBC", company),
         company = if_else(str_detect(company , "SHO|Showtime"), "Showtime", company),
         company = if_else(str_detect(company , "TBS"), "TBS", company))

```

# Primetime Emmys

### Programs
Outstanding Comedy Series
Outstanding Drama Series
Outstanding Limited or Anthology Series
Outstanding Competition Program
Outstanding Television Movie
Outstanding Variety Sketch Series
Outstanding Variety Talk Series

### Directing
Outstanding Directing for a Comedy Series
Outstanding Directing for a Drama Series
Outstanding Directing for a Limited or Anthology Series or Movie
Outstanding Directing for a Variety Series
### Writing
Outstanding Writing for a Comedy Series
Outstanding Writing for a Drama Series
Outstanding Writing for a Limited or Anthology Series or Movie
Outstanding Writing for a Variety Series

## Acting

### Lead actor
Outstanding Lead Actor in a Comedy Series
Outstanding Lead Actor in a Drama Series
Outstanding Lead Actor in a Limited or Anthology Series or Movie

### Lead actress
Outstanding Lead Actress in a Comedy Series
Outstanding Lead Actress in a Drama Series
Outstanding Lead Actress in a Limited or Anthology Series or Movie

### Supporting actor
Outstanding Supporting Actor in a Comedy Series
Outstanding Supporting Actor in a Drama Series
Outstanding Supporting Actor in a Limited or Anthology Series or Movie

### Supporting actress
Outstanding Supporting Actress in a Comedy Series
Outstanding Supporting Actress in a Drama Series
Outstanding Supporting Actress in a Limited or Anthology Series or Movie

```{r}

# Programs
outstanding_programs <- c(
"Outstanding Comedy Series",
"Outstanding Drama Series",
"Outstanding Limited Series",
"Outstanding Anthology Series",
"Outstanding Competition Program",
"Outstanding Television Movie",
"Outstanding Variety Sketch Series",
"Outstanding Variety Talk Series")

outstanding_lead_actor <- c(
"Outstanding Lead Actor In A Comedy Series",
"Outstanding Lead Actor In A Drama Series",
"Outstanding Lead Actor In A Limited Series Or Movie")

outstanding_lead_actress <- c(
"Outstanding Lead Actress In A Comedy Series",
"Outstanding Lead Actress In A Drama Series",
"Outstanding Lead Actress In A Limited Series Or Movie")

```

```{r}

emmys_primetime <-
emmys_cln %>%
  mutate(category = if_else(str_detect(category , paste(outstanding_programs , collapse = "|"))     , "Outstanding Programs", category),
         category = if_else(str_detect(category , "Outstanding Directing")                          , "Outstanding Directing", category),
         category = if_else(str_detect(category , "Outstanding Writing")                            , "Outstanding Writing", category),
         category = if_else(str_detect(category , paste(outstanding_lead_actor , collapse = "|"))   , "Outstanding Lead Actor", category),
         category = if_else(str_detect(category , paste(outstanding_lead_actress , collapse = "|")) , "Outstanding Lead Actress", category),
         category = if_else(str_detect(category , "Outstanding Supporting Actress")                 , "Outstanding Supporting Actress", category),
         category = if_else(str_detect(category , "Outstanding Supporting Actor")                   , "Outstanding Supporting Actor", category)) %>%
  filter(category %in% c("Outstanding Programs", 
                         "Outstanding Directing", 
                         "Outstanding Writing", 
                         "Outstanding Lead Actor", 
                         "Outstanding Lead Actress", 
                         "Outstanding Supporting Actress", 
                         "Outstanding Supporting Actor"))

```

```{r}

noms_10 <-
  emmys_primetime %>% 
  count(company) %>% 
  filter(n > 10) %>%
  select(company) %>%
  unlist(use.names = FALSE)

```

#### Sankey
```{r}
library(ggsankey)

emmy_sankey <-
  emmys_primetime %>%
  filter(company %in% noms_10) %>%
  select(company , category , type) %>% 
  make_long(company , category , type)

```

```{r}
library(showtext); showtext_auto()
font_add_google("Cutive Mono" , "cutive")
font_add_google("Cutive" , "cutive_bold")

emmy_sankey <-
emmy_sankey %>%
ggplot(aes(x = x, 
           next_x = next_x, 
           node = node, 
           next_node = next_node,
           fill = node)) +
  geom_sankey(flow.alpha = 0.5 , node.fill = "grey90" , node.color = "white" , width = 0.025 , show.legend = FALSE) +
  geom_sankey_text(aes(label = node), size = 5, color = "black" , family = "cutive" , show.legend = FALSE) +
  scale_fill_viridis_d(option = "turbo" , begin = 1 , end = 0.15) +
  labs(title = ,
       x = NULL) +
  theme_sankey() +
  theme(axis.text.x = element_blank(),
        plot.margin = margin(0,-6,0,-5.5,unit = "cm"))

```

```{r}
library(cowplot)

main_text <- 
"An Emmy Award, or simply Emmy, is a trophy presented at one of the numerous annual American\nevents or competitions that each recognize achievements in a particular sector of the television\nindustry. The Emmy is considered one of the four major entertainment awards in the United States,\nthe others being the Grammy (for music), Oscar (Academy Award) (for film), and Tony (for theatre).\nThe two events with the most media coverage are the Primetime Emmy Awards and the Daytime Emmy Awards,\nwhich recognize outstanding work in American primetime and daytime entertainment programming, respectively.\n\nThis graphic shows those companies which have had at least 10 nominations in the Primetime Emmy Awards\nmain categories for: Outstanding Programs, Writing, Directing, Lead Actors and Actresses, as well as Supporting\nActors and Actresses. The Primetime Emmy awards have always been a consistent category so this allows us to represent\na lot of data (since 1949) and how companies fed into these categories through nominations"

ggdraw(emmy_sankey) +
  draw_text("Emmy Awards\nand\nNominees" , family = "cutive_bold" , hjust = 1, size = 20, x = 0.90 , y = 0.9) +
  draw_text("Data\n\nGraphic" , family = "cutive" , hjust = 1, size = 12, x = 0.90 , y = 0.78) +
  draw_text("Emmys.com & Kaggle\n\n\n@NearAndDistant", family = "cutive_bold" , hjust = 1, size = 8 , x = 0.90 , y = 0.76) +
  draw_text(main_text , family = "cutive" , hjust = 1, size = 10, x = 0.90 , y = 0.15)

```

#### Saving
```{r}

ggsave(here::here("Week 39 : Emmy Awards/primetimeemmy.png"), dpi = 360, height = 10, width = 15)

```

