---
title: "Tidy Template"
author: "NearAndDistant"
date: "29/09/2021"
output: html_document
---

# Registered Nurses

***

This week we shall try {gt} and {gtExtra} to describe statistics of registered nurses across the US. We could integrate this view with a map on the graphic. See documentation, [here](https://gt.rstudio.com)

***

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

#### Setup & Import
```{r}
library(tidyverse)

# create driectory for new project
dir.create(here::here("Week 41 : Registered Nurses"))

# import data for project
nurses <-
read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-05/nurses.csv')

head(nurses)

```

#### Clean & Wrangle
```{r}

nurses_cln <-
  nurses %>%
  janitor::clean_names() %>%
  filter(year == 2020) %>%
  filter(state != "Puerto Rico") %>%
  mutate(pc_nurses = total_employed_rn,
         state = str_to_upper(state)) %>%
  select(state , annual_salary_median, annual_salary_avg) %>%
  pivot_longer(cols = -state, names_to = "metric") %>%
  arrange(state, metric) 


```

#### Analysis
```{r}
theme_set(theme_minimal())

nurses_cln %>%
  ggplot() +
  geom_col(aes(x = value, reorder(state, value), fill = metric), show.legend = FALSE) 

```

#### Palette
```{r}
library(showtext); showtext_auto()
font_add_google("Arya", "arya")
font_add_google("Spirax", "spirax")

background <- "#fff9fd"
palette    <- c("#b2c1c2" , "#e4b2ad")

```

#### Plotting

```{r}
# formatting

# Add 2 rows with 0 at the beginning of the data frame
nurses_med <- 
  nurses_cln %>% 
  filter(metric == "annual_salary_median") %>%
  add_row(value = c(0,0,0,0), state = c("za", "zb", "zc", "zd"), .after = 53)

nurses_avg <- 
  nurses_cln %>% 
  filter(metric == "annual_salary_avg") %>%
  add_row(value = c(0,0,0,0), state = c("za", "zb", "zc", "zd"), .after = 53) %>%
  mutate(id    = row_number(),
         angle = 90 - 360 * (id-0.5) / max(id),
         hjust = ifelse(angle < -90, 1, 0),
         angle = ifelse(angle < -90, angle +180, angle),
         color = ifelse(state %in% c("za", "zb", "zc", "zd"), "1", "0"),
         size  = ifelse(state %in% c("za", "zb", "zc", "zd"), "1", "0")) %>%
  select(-metric)

```

```{r, fig.align="center", fig.width=6, fig.height=6, fig.cap="Figure: Here is a really important caption."}

nurses_avg_med <- 
nurses_avg %>%
  ggplot(aes(x = state, y = value)) +
  geom_bar(aes(fill = "#b2c1c2"), alpha = 0.80 , stat = "identity", show.legend = FALSE, data = nurses_med) +
  geom_bar(aes(fill = "#e4b2ad"), alpha = 0, color = "#e4b2ad" , stat = "identity", show.legend = FALSE) +
  ylim(-25000, max(nurses_avg$value+20000)) +
  geom_text(aes(label = state, y = value + 8000 , angle = angle, hjust = hjust, color = color, size = size), 
            family = "arya" , show.legend = FALSE) + 
  scale_fill_manual(values = palette) +
  scale_color_manual(values = c("#000000", background)) +
  scale_size_manual(values  = c(3 , 0)) +
  coord_polar(start = 0, clip = "off") +
  theme_void() +
  theme(plot.background = element_rect(color = background , fill = background),
        panel.background = element_rect(color = background , fill = background),
        plot.margin = margin(0,0,0,0, unit = "cm"))

```

```{r}
# legend format

start <- nurses_avg$id[which(nurses_avg$state == "za")]
end   <- nurses_avg$id[which(nurses_avg$state == "zd")]

grid_manual <- 
  tibble(x    = c(end, end, end),
         xend = c(start, start, start),
         y = c(50000,90000,130000),
         labels = c("$40,000","$80,000","$120,000"))

plot_nurses_avg_med <- 
  nurses_avg_med +
  geom_segment(data = grid_manual,
               aes(x = x, xend = xend, y = y, yend = y),
               col = "grey50") +
  geom_text(data = grid_manual, aes(x = 54, y = y + 5000, label = labels),
            size = 2.5, col = "grey50", hjust = 0, angle = 15, family = "arya")

```

#### Panel
```{r}
library(cowplot)

# finalise panel with layered plotting or annotation
ggdraw(xlim = c(0.2, 0.8), ylim = c(0.05,0.975)) +
  draw_plot(plot_nurses_avg_med , x = 0.020 , y = 0.325, height = 0.6 , width = 0.6) +
  # main title
  draw_text("DIAGRAM OF THE SALARIES OF NURSES", x = 0.5 , y = 0.94, size = 22, family = "spirax") +
  draw_text("IN EACH STATE IN THE US", x = 0.5 , y = 0.90, size = 18, family = "arya") +
  draw_line(x = c(0.46,0.54) , y = c(0.873,0.873)) +
  draw_line(x = c(0.45,0.55) , y = c(0.87,0.87)) +
  # fig 1
  draw_text("1.\nJANUARY 1998 TO JANUARY 2020", x = 0.71 , y = 0.91, size = 12, family = "arya") +
  draw_line(x = c(0.68,0.74) , y = c(0.8875,0.8875)) +
  # fig 2
  draw_text("2.\nJANUARY 2020 TO JANUARY 2021", x = 0.32 , y = 0.91, size = 12, family = "arya") +
  draw_line(x = c(0.29,0.35) , y = c(0.8875,0.8875)) +
  theme(plot.background = element_rect(color = background , fill = background),
        panel.background = element_rect(color = background , fill = background))

```

#### Saving
```{r}

gtsave(here::here(".png"))

ggsave(here::here(".png"), dpi = 360, height = 10, width = 15)

```


R markdown allows to easily format your text. You can add [links](www.r-graph-gallery.com), write in **bold** or *italic*.

Further formatting can be found [here](https://holtzy.github.io/Pimp-my-rmd/)