---
title: "Tidy Template"
author: "NearAndDistant"
date: "29/09/2021"
output: html_document
---

# Registered Nurses

***

This week we shall try {gt} and {gtExtra} to describe statistics of registered nurses across the US. We could integrate this view with a map on the graphic. See documentation, [here](https://gt.rstudio.com)

***

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

#### Setup & Import
```{r}
library(tidyverse)

# create driectory for new project
dir.create(here::here("Week 41 : Registered Nurses"))

# import data for project
nurses <-
read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-05/nurses.csv')

head(nurses)

```

#### Clean & Wrangle
```{r}

nurses_cln <-
  nurses %>%
  janitor::clean_names() %>%
  filter(state != "Puerto Rico") %>%
  mutate(pc_nurses = total_employed_rn,
         state = str_to_upper(state))


```

#### Analysis
```{r}
theme_set(theme_minimal())

nurses_cln %>%
  ggplot() +
  geom_col(aes(x = value, reorder(state, value), fill = metric), show.legend = FALSE) 

```

#### Palette
```{r}
library(showtext); showtext_auto()
font_add_google("Arya", "arya")
font_add_google("Spirax", "spirax")
font_add_google("Uchen", "uchen")
font_add_google("Homemade Apple", "apple")
font_add_google("Charm", "charm")

background <- "#fff9fd"
palette    <- c("#b2c1c2" , "#e4b2ad")

```

#### Plotting Fig 1

```{r}

nurses_fig1 <- 
nurses_cln %>%
  filter(year != 2021) %>%
  select(state , year , annual_salary_median, annual_salary_avg) %>%
  group_by(year) %>%
  summarise(annual_salary_avg    = mean(annual_salary_avg, na.rm = TRUE),
            annual_salary_median = median(annual_salary_median, na.rm = TRUE)) %>%
  pivot_longer(cols = -year, names_to = "metric")

```

```{r}

# Add 2 rows with 0 at the beginning of the data frame
fig1_med <- 
  nurses_fig1 %>% 
  filter(metric == "annual_salary_median") %>%
  add_row(value = c(0,0,0,0,0,0,0), year = c(2021:2027))

fig1_avg <- 
  nurses_fig1 %>% 
  filter(metric == "annual_salary_avg") %>%
  add_row(value = c(0,0,0,0,0,0,0), year = c(2021:2027)) %>%
  mutate(id    = row_number(),
         angle = 360 - 360 * (id - 0.5) / max(id),
         hjust = ifelse(angle < -90, 1, 0),
         angle = ifelse(angle < -90, angle +180, angle),
         color = ifelse(year %in% c(2021:2027), "1", "0"),
         size  = ifelse(year %in% c(2021:2027), "1", "0")) %>%
  select(-metric)

```

```{r, fig.align="center", fig.width=6, fig.height=6, fig.cap="Figure: Here is a really important caption."}

fig1_avg_med <- 
fig1_avg %>%
  ggplot(aes(x = year, y = value)) +
  geom_bar(aes(fill = "#b2c1c2"), alpha = 0.80 , stat = "identity", show.legend = FALSE, data = fig1_med) +
  geom_bar(aes(fill = "#e4b2ad"), alpha = 0, color = "#e4b2ad" , stat = "identity", show.legend = FALSE) +
  ylim(-25000, max(fig1_avg$value+20000)) +
  geom_text(aes(label = year, y = value + 4000, angle = angle, color = color, size = size), 
            family = "arya", fontface = "bold"  , show.legend = FALSE) + 
  scale_fill_manual(values = palette) +
  scale_color_manual(values = c("#000000", background)) +
  scale_size_manual(values  = c(4 , 0)) +
  coord_polar(start = 0, clip = "off") +
  theme_void() +
  theme(plot.background = element_rect(color = background , fill = background),
        panel.background = element_rect(color = background , fill = background),
        plot.margin = margin(0,0,0,0, unit = "cm"))

```

```{r}
# legend format fig 1

start1 <- 2021
end1   <- max(fig1_avg$year)

grid_fig1 <- 
  tibble(x      = c(start1, start1, start1),
         xend   = c(end1  , end1  , end1  ),
         y      = c(25000 , 50000 , 75000 ),
         labels = c("$25,000","$50,000","$75,000"))

plot_fig1 <- 
  fig1_avg_med +
  geom_segment(data = grid_fig1, col = "grey50", 
               aes(x = x, xend = xend, y = y, yend = y)) +
  geom_text(data = grid_fig1, aes(x = max(fig1_avg$year), y = y + 5000, label = labels),
            size = 3, col = "grey50", fontface = "bold" , hjust = 1, angle = 0, family = "arya")

```

#### Plotting Fig 2

```{r}

nurses_fig2 <- 
nurses_cln %>%
  filter(year == 2020) %>%
  select(state , annual_salary_median, annual_salary_avg) %>%
  pivot_longer(cols = -state, names_to = "metric") %>%
  arrange(state, metric) 

```

```{r}
# formatting

# Add 2 rows with 0 at the beginning of the data frame
nurses_med <- 
  nurses_fig2 %>% 
  filter(metric == "annual_salary_median") %>%
  add_row(value = c(0,0,0,0), state = c("za", "zb", "zc", "zd"), .after = 53)

nurses_avg <- 
  nurses_fig2 %>% 
  filter(metric == "annual_salary_avg") %>%
  add_row(value = c(0,0,0,0), state = c("za", "zb", "zc", "zd"), .after = 53) %>%
  mutate(id    = row_number(),
         angle = 90 - 360 * (id-0.5) / max(id),
         hjust = ifelse(angle < -90, 1, 0),
         angle = ifelse(angle < -90, angle +180, angle),
         color = ifelse(state %in% c("za", "zb", "zc", "zd"), "1", "0"),
         size  = ifelse(state %in% c("za", "zb", "zc", "zd"), "1", "0")) %>%
  select(-metric)

```

```{r, fig.align="center", fig.width=6, fig.height=6, fig.cap="Figure: Here is a really important caption."}

nurses_avg_med <- 
nurses_avg %>%
  ggplot(aes(x = state, y = value)) +
  geom_bar(aes(fill = "#b2c1c2"), alpha = 0.80 , stat = "identity", show.legend = FALSE, data = nurses_med) +
  geom_bar(aes(fill = "#e4b2ad"), alpha = 0, color = "#e4b2ad" , stat = "identity", show.legend = FALSE) +
  ylim(-25000, max(nurses_avg$value+20000)) +
  geom_text(aes(label = state, y = value + 8000 , angle = angle, hjust = hjust, color = color, size = size), 
            family = "arya" , fontface = "bold" , show.legend = FALSE) + 
  scale_fill_manual(values = palette) +
  scale_color_manual(values = c("#000000", background)) +
  scale_size_manual(values  = c(3 , 0)) +
  coord_polar(start = 0, clip = "off") +
  theme_void() +
  theme(plot.background = element_rect(color = background , fill = background),
        panel.background = element_rect(color = background , fill = background),
        plot.margin = margin(0,0,0,0, unit = "cm"))

```

```{r}
# legend format

start <- nurses_avg$id[which(nurses_avg$state == "za")]
end   <- nurses_avg$id[which(nurses_avg$state == "zd")]

grid_manual <- 
  tibble(x    = c(end, end, end),
         xend = c(start, start, start),
         y = c(50000,90000,130000),
         labels = c("$40,000","$80,000","$120,000"))

plot_fig2 <- 
  nurses_avg_med +
  geom_segment(data = grid_manual,
               aes(x = x, xend = xend, y = y, yend = y),
               col = "grey50") +
  geom_text(data = grid_manual, aes(x = 54, y = y + 5000, label = labels),
            size = 2.5, col = "grey50", hjust = 0, angle = 15, family = "arya", fontface = "bold")

```

#### Panel

```{r}

text <-"The Areas of the blue and red outline are measured from the centreas the\n   common vertex.\nThe blue wedges measured from the centre of the circle represent area for area the\n   median salary for nurses in each year (fig 1)\n   and each state in 2021 (fig 2); & the red outline measures the mean.\nWe can see that across years there has been a close relationship between\n   the median and mean for salaries. When we look at a state level we can see the\n   sometimes large discrepancies in salaries although the close relationship between\n   median and mean still hold.\nHawaii is an exception to the general rule where median salary is larger than\n    mean, we could expect that some individuals in Hawaii earn a lot less than other,\n    dragging the central tendancy of the mean downwards."

```

```{r}
library(cowplot)

# finalise panel with layered plotting or annotation
ggdraw(xlim = c(0.23, 0.8), ylim = c(0.10,0.975)) +
  # plot 1
  draw_plot(plot_fig1 , x = 0.15 , y = 0.01, height = 1 , width = 1) +
  # plot 2
  draw_plot(plot_fig2 , x = 0.07 , y = 0.40, height = 0.5 , width = 0.5) +
  # main title
  draw_text("DIAGRAM            SALARIES       NURSES", x = 0.5 , y = 0.94, size = 22, family = "spirax") +
  draw_text("      OF THE                                                OF", x = 0.5 , y = 0.935,size = 12, family = "spirax") +
  draw_text("IN THE UNITED STATES", x = 0.5 , y = 0.90, size = 18, family = "arya") +
  draw_line(x = c(0.46,0.54) , y = c(0.873,0.873)) +
  draw_line(x = c(0.45,0.55) , y = c(0.87,0.87)) +
  # fig 1
  draw_text("1.\nJANUARY 1998 TO JANUARY 2020", x = 0.68 , y = 0.91, size = 10, family = "uchen") +
  draw_line(x = c(0.65,0.71) , y = c(0.8875,0.8875)) +
  # fig 2
  draw_text("2.\nJANUARY 2020 TO JANUARY 2021", x = 0.32 , y = 0.91, size = 10, family = "uchen") +
  draw_line(x = c(0.29,0.35) , y = c(0.8875,0.8875)) +
  # connection
  annotate("segment", x = 0.5075, y = 0.5425, xend = 0.38, yend = 0.42, linetype = "dashed") +
  annotate("segment", x = 0.38, y = 0.42, xend = 0.32, yend = 0.46, linetype = "dashed") +
  # text
  draw_text(text , x = 0.25, y = 0.26 , hjust = 0, size = 13 , family = "charm") +
  draw_text("Data: U.S. BUREAU OF LABOR STATISTICS | Graphic: @NearAndDistant",
            x = 0.65 , y = 0.125 , size = 10 , family = "arya") +
  theme(plot.background = element_rect(color = background , fill = background),
        panel.background = element_rect(color = background , fill = background))

```

#### Saving
```{r}

ggsave(here::here("Week 41 : Registered Nurses/nightingale.png"), dpi = 360, height = 10, width = 15)

```


R markdown allows to easily format your text. You can add [links](www.r-graph-gallery.com), write in **bold** or *italic*.

Further formatting can be found [here](https://holtzy.github.io/Pimp-my-rmd/)