---
title: '#30DayMapChallenge'
author: "NearAndDistant"
date: "04/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 30 Day Map Challenge

# Day 01 : Points

```{r}



```


# Day 02 : Lines

```{r}

```


# Day 03 : Polygons

```{r}

```


# Day 04 : Hexagons

```{r}

```


# Day 05 : Open Street Map

```{r}

```


# Day 06 : Red

```{r}

```


# Day 07 : Green

```{r}

```


# Day 08 : Blue

```{r}

```


# Day 09 : Monochrome

```{r}

```


# Day 10 : Raster

```{r}

```


# Day 11 : 3D

```{r}

```


# Day 12 : Population

```{r}

```


# Day 13 : Natural Earth

```{r}

```


# Day 14 : New Tool

```{r}

```


# Day 15 : No Computer Map

```{r}

```


# Day 16 : Urban / Rural

```{r}

```


# Day 17 : Land

```{r}

```


# Day 18 : Water

```{r}

```


# Day 19 : Island

```{r}

```


# Day 20 : Movement

```{r}

```


# Day 21 : Elevation

```{r}

```


# Day 22 : Boundaries

```{r}

```


# Day 23 : GHSL

```{r}

```


# Day 24 : Historical

```{r}

```


# Day 25 : Interactive

```{r}

```


# Day 26 : Chloropleth

```{r}

```


# Day 27 : Heatmap

```{r}

```


# Day 28 : Earth in not Flat

```{r}

```


# Day 29 : NULL

```{r}

```


# Day 30 : Metamapping

```{r}

```


